### 2022-04-07 README.Rmd
small enhancement

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Estimation of Population Density for the prehistoric Swiss Plateau using a Hierarchical Bayesian Poisson Regression Model"
```

# bayesian.demographic.reconstruction.2022

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/MartinHinz/bayesian.demographic.reconstruction.2021/master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### How to cite

Please cite this compendium as:

> Authors, (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

## Contents

The **repository** consists of:

  - [:file\_folder: manuscript](/manuscript): R Markdown source document
    for manuscript.
  - [:file\_folder: paper](/paper): A rendered version, of the manuscript
    `paper.docx` and `paper.pdf`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: analysis](/analysis): R Markdown source document
    for the actual analysis. Includes code to reproduce the figures and tables
    generated by the analysis. You can look at the [Github-Markdown Version](/analysis/analysis.md)
    for a rendered version of the analytical run incl. figures.
  - [:file\_folder: code](/code): Functions used in the analysis.
  - [:file\_folder: data](/data): Data used in the analysis.
  - [:file\_folder: figures](/figures): Plots and other illustrations

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.